\part{Foundations}
\label{part:background}

\chapter{Background}
\label{ch:background}

This chapter presents the terminology used in the rest of the thesis. We discuss the evolution of the Web into the Semantic Web, and how it became popular as Linked Data and through the Linking Open Data project. We describe some of the core concepts and technologies of the Semantic Web, like ontologies and RDF. We also present a list of Personal Information Management studies, and we define the meaning with which the term ``personal information'' is used in this thesis.

\input{chapters/background/semweb}

\input{chapters/background/pim}

\section{Conclusion}

In this chapter we presented the terminology that is used in the thesis, both from the area of the Semantic Web and that of Personal Information Management. We gave an overview of the Semantic Web, its evolution from the document Web, and its core technologies. 
We have also described what ``personal information'' refers to in the context of Personal Information Management, and what are the research areas, and challenges of the domain. 

We continue to show in the next chapter how the semantic technologies have been applied on the desktop and to PIM, to create Semantic Desktop systems. These systems are apply the most basic elements of semantic technologies, the categorisation of entities into classes and the explicit specification of links, to representing the mental models with the help of computers. 

\chapter{The Semantic Desktop}
\label{ch:sd}

\begin{flushright}
 \textit{Based on ``Knowledge Management on the Desktop'' \cite{Dragan2012}\\published at the 18th International Conference on Knowledge Engineering and Knowledge Management (EKAW2012)}
\end{flushright}

Having defined the semantic technologies used on the Web, we continue in this chapter to show how they have been applied on the desktop and to Personal Information Management, to create the Semantic Desktop. We provide an in-depth overview of existing Semantic Desktop features and architectures, since our work is grounded in the framework given by the Semantic Desktop, enhancing and complementing its features. 

\input{chapters/background/historicalsd}

\input{chapters/background/researchsd}
