\part{Prelude}
\label{part:prelude}

\chapter{Introduction}
\label{ch:introduction}

We accumulate massive amounts of information on our computers and other devices, so much that it becomes daunting and sometimes useless because it is impossible to remember it all: emails, appointments, tasks, documents, people, conversations, etc. In traditional desktop architectures this information is created and stored in various applications, each an isolated island of data, with its own storage and its own format, unaware of related and relevant data in other applications. But in our mind this information is not separate, rather everything is connected by implicit links. When we use information we do not work with just one detached slice of it, but we mentally follow connections to other pieces of information from multiple sources. Information is inherently interlinked and used in an integrated fashion.

The Semantic Desktop emerged as a solution to the data fragmentation and disconnection problem. It provides a foundation and a framework based on Semantic Web standards and technologies, for interlinking information on the desktop. Interlinked information is easier to manage and organise, and because it mimics users' mental models, it is easier to find when required, thus also offering a solution to the information overload problem. The Semantic Desktop overcomes limitations of conventional desktops by using a common representation for the data, common vocabularies to describe it, and a central place to store it. It creates a network of linked desktop data, in a standardised format, which is accessible to all applications.

Several implementations of Semantic Desktops exist, with varied function sets and capabilities. NEPOMUK is a project undertaking the task of creating a blueprint for the Semantic Desktop. 
The work presented in this thesis is grounded and built upon the framework provided by Nepomuk-KDE, a branch of the NEPOMUK project. 

The most common use case for the Semantic Desktop is Personal Information Management (PIM). However, having a relatively stable framework for the Semantic Desktop does not bring the search for improved PIM to an end, rather it uncovers a path to a solution. There is plenty of work ahead, exploiting and improving the framework based on experience, research and testing. This thesis focuses on providing new and improved means of creating and maintaining the network of interconnected personal semantic data, on the desktop and beyond.

\input{chapters/introduction/problem}

\input{chapters/introduction/contribution}

\input{chapters/introduction/structure}
