\chapter{Additional Extensions and Applications}
\label{ch:mischelperapps}

In this chapter we describe several applications developed in our SmILE\footnote{\url{http://smile.deri.ie}} group at DERI, and to which we contributed. They are extensions to the work presented in the preceding pages, providing or serving as test use cases and validation scenarios.

We start with a list of extensions to SemNotes' functionality, which employ Natural Language Processing techniques. They focus on specific use cases of note-taking, like ontology engineering, meeting minutes or status reports; as well as text analytics tailored to short text.

The work presented in Chapter \ref{ch:sdwod} was initially motivated by the increasingly difficult task of managing publications on the desktop and on the Web, and the need of connecting the information available in both worlds. \emph{Sclippy} is a tool created to avail of the rich information available on the Web about publications, and bring it to the user's desktop. 
The algorithm described in Chapter \ref{ch:sdwod} is a generalisation of the algorithms used in Sclippy\footnote{\url{http://smile.deri.ie/projects/sclippy}}, extending the scope from publications to any type of desktop resources.

We continue by presenting Konduit, a desktop-based platform for visual scripting with RDF data. Based on the idea of the Semantic Desktop, non-technical users can create, manipulate and mash-up RDF data with Konduit, and thus generate simple applications or workflows, which are aimed to simplify their everyday work by automating repetitive tasks. The platform allows to combine data from both the Web and the desktop and integrate it with existing desktop functionality.

\input{chapters/core/misc/nlpcnl}

\input{chapters/core/misc/sclippy}

\input{chapters/core/misc/konduit}
