\chapter
{Creating and Interlinking Semantic Data on the Desktop with SemNotes}
\label{ch:semnotes}

\begin{flushright}
 \textit{Based on ``The Semantic Desktop at Work: Interlinking Notes'' \cite{Dragan2011a}\\published at the 7th International Conference on Semantic Systems (I-SEMANTICS~2011)}
\end{flushright}

The Semantic Desktop has been proposed as a solution to the ever growing problem of information overload on our computers. It provides the foundations necessary to integrate and manage personal information. However, the challenge of designing and realising semantic applications that use this infrastructure still remains. 
In this chapter, we present SemNotes\footnote{\url{http://smile.deri.ie/projects/semn}}, a semantic note-taking tool for the Nepomuk-KDE Semantic Desktop, as a tool for creating new semantic information on the desktop and integrating it seamlessly in the existing network of linked desktop data.
SemNotes provides a real-world, functional use case for fully exploiting the capabilities of the Semantic Desktop: interlinking, organisation and management of personal information, improved search and browsing. 
Furthermore, it represents our solution to a set of identified generic challenges for applications on the Semantic Desktop, as described by the first research question in Section \ref{sub:question}.
We describe a task-based user evaluation comparing SemNotes with a conventional note-taking tool. The results show that complex searches on interlinked information created with SemNotes are significantly faster, with little or no extra effort required from the users. 

\input{chapters/core/semnotes/introduction}

\input{chapters/core/semnotes/design}

\input{chapters/core/semnotes/implementation}

\input{chapters/core/semnotes/evaluation}

\input{chapters/core/semnotes/related}

\section{Conclusion}

In this chapter, we presented a solution to the challenges of designing applications for the Semantic Desktop. We described one such application, SemNotes, a semantic note-taking tool for the Nepomuk-KDE Semantic Desktop. It provides a real-world, functional use case for fully exploiting the capabilities of the Semantic Desktop: interlinking, organisation and management of personal information, efficient search and browsing. It supports the entire life-cycle of the semantic data represented by notes, with emphasis on the creation of links between the new data and the existing network of linked information on the desktop. 

Through SemNotes we present a possible solution to each of the challenges presented in the begining of the chapter. To the first question of creating semantic data, we describe how we create new semantic notes, using the existing vocabularies and creating links to the existing Semantic Desktop data available. To the second question of designing the human computer interaction, we describe the efforts towards an easy to use yet rich user interface design for SemNotes. With the help of usability experts and user testing the interface went so far through three iterations, each described in this chapter.

For the third challenge, that of evaluating a semantic tool, we described a task-based user evaluation comparing SemNotes to the popular note-taking application Evernote. The results show that the extra effort (measured in time spent) required for the annotation of notes with links to related resources is not significant. However, the benefit (measured in time saved) is significant for one of two complex search tasks. 

SemNotes embodies a simple and user-friendly way of generating new semantic data on the desktop, and integrating it with the already existing data. However, the personal information that the users work with is not restricted to the desktop. Further, semantic data exists in large quantities outside of the desktop, on the Web and in organisational repositories. Thus, in the next chapter we continue by describing how the linked information from the desktop can be meaningfully and safely connected to the outside world.
